---
title: "Sentimentanalysis_posts"
author: "Jose Hernandez"
date: "2026-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview of lesson

In this lesson, we will:
- Load a CSV file of short text posts
- Clean and tokenize the text
- Perform sentiment analysis using the NRC lexicon
- Visualize emotions with barplots and word clouds


```{r}
library(syuzhet)
library(RColorBrewer)
library(wordcloud)
library(readr)
library(dplyr)

```

### Loading data

First we will load our posts.


```{r}
data <- read_csv("sentiment_statements_unlabeled.csv")

# View data
head(data)
```

### Text Processing

Now we combine the text and we tokenize it by word. 

```{r}
# joining the posts
text_string <- paste(data$statement_text, collapse = " ")

text_words <- get_tokens(text_string)
head(text_words)


```

We could also tokenize it by sentence instead

```{r}
sentence_vector <- get_sentences(text_string)
head(sentence_vector)

```

### Sentiment analysis
```{r}

sentiment_scores <- get_nrc_sentiment(text_words, lang = "english")

head(sentiment_scores)
```
Given how large our dataset is it is important thbat we have a way to view it in context either by a statistical summary or a plot.

```{r}
summary(sentiment_scores)
```


### 
```{r}
barplot(
  colSums(prop.table(sentiment_scores[, 1:8])),
  space = 0.2,
  horiz = FALSE,
  las = 1,
  cex.names = 0.8,
  col = brewer.pal(n = 8, name = "Set3"),
  main = "Emotion Distribution of Posts",
  xlab = "Emotions",
  ylab = "Proportion"
)

barplot(
  colSums(prop.table(sentiment_scores[, 9:10])),
  space = 0.2,
  horiz = FALSE,
  las = 1,
  cex.names = 0.8,
  col = brewer.pal(n = 8, name = "Set3"),
  main = "Positive vs Negative Distribution of Posts",
  xlab = "Categories",
  ylab = "Proportion"
)

```


# Parallelization 

parallelization allow us to assign our tasks to multiple cores to get the work done quicker. But first we need to learn how to reus our code.

### Reusing our code -Creating a function 

To do this, we have to create a function that runs this sentiment analysis once we call it.
```{r}

run_sentiment <- function(words_subset) {
  get_nrc_sentiment(words_subset, lang = "english")
}

```



### Distributing the work

Having the function we can now split our dataset.
```{r}
words_part1 <- text_words[1:1500]
words_part2 <- text_words[1501:3029]
```



### Applying a function to our work

Now instead of writing our sentiment analysis code block again we can run the function and apply it to bboth of our datasets. 

```{r}
baseline_time <- system.time({
  results_linear <- lapply(
      list(words_part1, words_part2),
      run_sentiment
  )
})
  
print(baseline_time)
```



### Applying the fucntion in parallel

```{r}
library(parallel)
```

Before creating our parallelization lets first see how many cores we have available to use.

```{r}
num_cores <- detectCores() 
print(num_cores)


allocated_cores <- as.numeric(Sys.getenv("SLURM_JOB_CPUS_PER_NODE", unset = 1))
print(allocated_cores)


```

```{r}
parallel_time <- system.time({
  
  results_parallel <- mclapply(
    list(words_part1, words_part2),
    run_sentiment,
    mc.cores = 2
  )
  
  # Combine results back together
  sentiment_scores_parallel <- do.call(rbind, results_parallel)
})

print(parallel_time)


```


